\LARGE \bf Approximately Orchestrated Routing and Transportation Analyzer:\\ Large-scale Traffic Simulation for Autonomous Vehicles

### 2012-08-29 papers/itsc_2012/submitted2.tex

Autonomous vehicle technology has made tremendous progress in the last decade.
In 2007, six of the competing teams completed the 96km course set for the DARPA
Urban Challenge \cite{spectrumUrbanChallenge2007}. They did so while
obeying the traffic laws followed by human drivers, navigating along with other
moving vehicles, and following correct intersection precedence order. Since
then, Google's driverless cars have clocked more than 250,000km on public roads
in urban California, USA \cite{tedThrun2011}. In 2010, researchers from the
University of Parma successfully completed an autonomous
\textit{intercontinental} run from Parma, Italy to Shanghai, China
\cite{cnnVislab2010}. The successful completion of all these milestones
suggests that autonomous cars are here to stay, and are ever closer to becoming
commercially available. With the arrival of autonomous cars, it also becomes
possible to optimize traffic in ways not possible for human drivers.

@@

Another popular open source traffic simulator that can use OSM data is SUMO
\cite{SUMO2011}, which shares many similar goals with AORTA. SUMO has been used to
study the effect of automated transportation systems, route planning of
individual vehicles, as well as dynamically adapting traffic signal policies to
increase traffic efficiency. While SUMO has a flexible system for managing
traditional traffic signals, AORTA's general intersection policies are designed
with autonomous vehicles in mind. These policies can use traditional schemes
such as traffic signals or stop-sign based precedence, or more efficient methods
designed specifically for autonomous vehicles \cite{JAIR08-dresner}. In
addition, AORTA's core implementation, written in Scala \cite{scala}, differs
from that of SUMO, written in C++, in that it uses a higher-level language.

@@

Currently agents cannot change lanes. This causes a second potential issue:
gridlock.  Each agent maintains a safe following distance from the next. If a
lane is filled, then an agent may be forced to stop in an intersection mid-turn
and block other traffic. The default behavior prevents this by refusing to start
a turn before the lookahead engine guarantees no agent will trigger a premature
stop.

However, with very high volumes of traffic, it is possible for waiting agents to
accumulate and fill lanes to their capacity. When this happens in a circular
manner so that the agent at the front of each lane is waiting to perform a turn
into another full lane with a similar agent at the front, gridlock
\cite{AAAI11-au} occurs: no agent in the system will make progress unless an
agent at the front decides to pick a different turn. AORTA has the ability to
find such cycles.

@@

The experiment simulates one hour of traffic in two cities: a
9$\times$11km$^{2}$ slice of downtown Austin, Texas\footnote{bounded by the
Mopac, Springdale Road, Oltorf Street and Koenig Lane}, and a
15$\times$13km$^{2}$ slice of downtown Houston, Texas\footnote{bounded by Ella
Boulevard, Eastex Freeway, West Alabama Street and Crosstimbers Street}.  The
step duration is fixed at 0.1 seconds. While the intersection policies vary,
the agents spawned and their routes remain the same.  A generator creates one
new agent every 1 simulation second, with a uniformly random starting position
and goal.

>>>>

Autonomous vehicle technology has made tremendous progress in the last decade.
In 2007, six of the competing teams completed the 96 km course set for the DARPA
Urban Challenge \cite{spectrumUrbanChallenge2007}. They did so while
obeying the traffic laws followed by human drivers, navigating along with other
moving vehicles, and following correct intersection precedence order. Since
then, Google's driverless cars have clocked more than 250,000 km on public roads
in urban California, USA \cite{tedThrun2011}. In 2010, researchers from the
University of Parma successfully completed an autonomous
\textit{intercontinental} run from Parma, Italy to Shanghai, China
\cite{cnnVislab2010}. The successful completion of all these milestones
suggests that autonomous cars are here to stay, and are ever closer to becoming
commercially available. With the arrival of autonomous cars, it also becomes
possible to optimize traffic in ways not possible for human drivers.

@@

Another popular open source traffic simulator that can use OSM data is SUMO
\cite{SUMO2011}, which shares many similar goals with AORTA. SUMO has been used to
study the effect of automated transportation systems, route planning of
individual vehicles, and dynamically adapting traffic signal policies to
increase traffic efficiency. While SUMO has a flexible system for managing
traditional traffic signals, AORTA's general intersection policies are designed
with autonomous vehicles in mind. These policies can use traditional schemes
such as traffic signals or stop-sign based precedence, or more efficient methods
designed specifically for autonomous vehicles \cite{JAIR08-dresner}. In
addition, AORTA's core implementation, written in Scala \cite{scala}, differs
from that of SUMO, written in C++, in that it uses a higher-level language.

@@

In the currently released version of the simulator, the ability for agents to
change lanes has not yet been implemented. As a result, gridlock potentially
occurs. Each agent maintains a safe following distance from the next. If a lane
is filled, then an agent may be forced to stop in an intersection mid-turn and
block other traffic. The default behavior prevents this by refusing to start a
turn before the lookahead engine guarantees no agent will trigger a premature
stop.

However, with very high volumes of traffic, it is possible for waiting agents to
accumulate and fill lanes to their capacity. When this happens in a circular
manner so that the agent at the front of each lane is waiting to perform a turn
into another full lane with a similar agent at the front, gridlock
\cite{AAAI11-au} occurs: no agent in the system will make progress unless an
agent at the front decides to pick a different turn. AORTA has the ability to
detect such cyclic dependencies.

@@

The experiment simulates one hour of traffic in two cities: a
9$\times$11 km$^{2}$ slice of downtown Austin, Texas\footnote{bounded by the
Mopac, Springdale Road, Oltorf Street and Koenig Lane}, and a
15$\times$13 km$^{2}$ slice of downtown Houston, Texas\footnote{bounded by Ella
Boulevard, Eastex Freeway, West Alabama Street and Crosstimbers Street}.  The
step duration is fixed at 0.1 seconds. While the intersection policies vary,
the agents spawned and their routes remain the same.  A generator creates one
new agent every 1 simulation second, with a uniformly random starting position
and goal.
